Variational Monte Carlo on bosonic systems

### 2018-03-06 project_1/article/article.tex
Add missing boy
Add references
	The placement of the citations is completely arbitrary.
Add theory on scaling the system

\section{Introduction}
    We will in this project study the Variational Monte Carlo (VMC) method, and
    use it to evaluate the ground state energy of a trapped, hard sphere Bose
    gas.

@@

@@

    \subsection{Statistical Analysis}
        If the results of the metropolis sampling were completely uncorrelated,
        it would be enough to compute the standard deviation in a familiar way,
        \begin{equation}
            \sigma = \sqrt{\frac{1}{N}(\ev{E^2_L} - \ev{E_L}^2},
            \label{eq:standard_deviation}
        \end{equation}
        where $N$ is the number of samples, or Monte-Carlo cycles, in the experiment.
        However, it is reasonable to assume that the data we are dealing with in this
        study is liable to suffer from \emph{autocorrelation} and \autoref{eq:standard_deviation}
        does not hold. The prevailing
        definition of autocorrelation in a data stream or signal is correlation
        between a delay of the signal and the original signal. One would be interested to find the
        delay, or lag, in the signal at which the "self-correlation" is highest. We refer to this
        spacing as $d$, and define the following correlation function,
        \begin{equation}
            f_d = \frac{1}{n -d} \sum_{k=1}^{n-d} (x_k - \bar{x}_n)(x_{k+d} - \bar{x}_n).
            \label{eq:correlation_f_d}
        \end{equation}
        The keen reader would have noticed that the function $f_d$ in \autoref{eq:correlation_f_d}
        would be equal to the sample variance for $d=0$. We can now define the \emph{autocorrelation function}
        \begin{equation}
            \kappa_d = \frac{f_d}{\text{Var}(x)},
            \label{eq:autocorrelation_function}
        \end{equation}
        which is equal to $1$ if the data exhibits no autocorrealtions, equating to $d=0$. From the autocorrelation
        function (\ref{eq:autocorrelation_function}) we in turn define the \emph{autocorrelation time},
        \begin{equation}
            \tau = 1 + 2 \sum_{d=1}^{n-1},
        \end{equation}
        notice that the autocorrelation time is $1$ for a correlation free experiment.

@@

\end{document}

>>>>

\section{Introduction}
    We will in this project study the Variational Monte Carlo (VMC) method, and
    use it to evaluate the ground state energy of a trapped, hard sphere Bose
    gas. \cite{dubois2001bose} \cite{nilsen2005vortices}

@@

    \subsection{Scaling the system}
        We now introduce a scaled distance $\vf{r}' = \vf{r}/a_{\text{ho}}$
        \cite{dubois2001bose}, where
        \begin{align}
            a_{\text{ho}}
            &=
            \sqrt{\frac{\hbar}{m\omega}}.
        \end{align}
        We can then rewrite the Hamiltonian for the elliptic potential in terms
        of this new distance. By doing a variable substitution for each
        direction in the Laplace operator we get
        \begin{align}
            \nabla_k^2
            &=
            \frac{1}{a_{\text{ho}}^2}{\nabla_k'}^2.
        \end{align}

        Looking at the one-body part of the Hamiltonian we get
        \begin{align}
            H
            &=
            \sum_{k = 1}^N
            \brac{
                -\frac{\hbar^2}{2m}\nabla_k^2
                + \half m\brak{
                    \omega^2(x_k^2 + y_k^2)
                    + \omega_z^2 z_k^2
                }
            }
            \\
            &=
            \frac{\hbar\omega}{2}
            \sum_{k = 1}^N
            \brac{
                -{\nabla_k'}^2
                +
                \brak{
                    {x_k'}^2 + {y_k'}^2 + \lambda^2{z_k'}^2
                }
            },
        \end{align}
        where we have introduced the dimensionless frequency $\lambda =
        \omega_z/\omega$. The single particle functions also gets a scaling
        factor
        \begin{align}
            \phi_k
            &=
            \exp\brak{
                -\frac{\alpha\hbar}{m\omega}
                ({x_k'}^2 + {y_k'}^2 + \beta^2{z_k'}^2)
            }
            \equiv
            \phi(\vf{r}_k').
        \end{align}
        The correlation wavefunction and the interaction potential remain the
        same, but using the scaled hard core sphere radius $a/a_{\text{ho}} =
        0.0043$. For brevity we will remove the ticks on the variables.

@@

    \subsection{Statistical Analysis}
        If the results of the metropolis sampling were completely uncorrelated,
        it would be enough to compute the standard deviation in a familiar way,
        \begin{equation}
            \sigma = \sqrt{\frac{1}{N}(\ev{E^2_L} - \ev{E_L}^2},
            \label{eq:standard_deviation}
        \end{equation}
        where $N$ is the number of samples, or Monte-Carlo cycles, in the experiment.
        However, it is reasonable to assume that the data we are dealing with in this
        study is liable to suffer from \emph{autocorrelation} and \autoref{eq:standard_deviation}
        does not hold. The prevailing
        definition of autocorrelation in a data stream or signal is correlation
        between a delay of the signal and the original signal. One would be interested to find the
        delay, or lag, in the signal at which the "self-correlation" is highest. We refer to this
        spacing as $d$, and define the following correlation function,
        \begin{equation}
            f_d = \frac{1}{n -d} \sum_{k=1}^{n-d} (x_k - \bar{x}_n)(x_{k+d} - \bar{x}_n).
            \label{eq:correlation_f_d}
        \end{equation}
        The keen reader would have noticed that the function $f_d$ in \autoref{eq:correlation_f_d}
        would be equal to the sample variance for $d=0$. We can now define the \emph{autocorrelation function}
        \begin{equation}
            \kappa_d = \frac{f_d}{\text{Var}(x)},
            \label{eq:autocorrelation_function}
        \end{equation}
        which is equal to $1$ if the data exhibits no autocorrealtions, equating to $d=0$. From the autocorrelation
        function (\ref{eq:autocorrelation_function}) we in turn define the \emph{autocorrelation time},
        \begin{equation}
            \tau = 1 + 2 \sum_{d=1}^{n-1}\kappa_d,
        \end{equation}
        notice that the autocorrelation time is $1$ for a correlation free experiment.

@@

\bibliography{references}
\end{document}
